\section{Introduction}

The convergence of large language models with high-performance computing is transforming scientific workflows. Autonomous agents now steer simulations in real time, orchestrate multi-node ML training pipelines, and analyze petabytes of experimental data without human intervention. These AI agents promise to dramatically accelerate discovery, but they also introduce a threat category that HPC security has never confronted.

The core problem is not that AI agents are untrustworthy. The problem is that they are too trustworthy, specifically, they trust the wrong inputs. An agent authorized to process genomics data for Project X will faithfully execute whatever instructions it receives, including instructions embedded within that data by an attacker. A prompt injection payload hidden in a \textit{FITS} file header, a malicious command left in a shared \textit{/tmp} directory by a co-located job, or a compromised tool that returns hidden instructions during normal operation represent concrete attack vectors. These are not theoretical concerns; they are the inevitable consequence of deploying instruction-following systems into environments where untrusted data is the norm.

This paper identifies and formalizes a threat that, to our knowledge, has not been studied in prior work: \textit{the hijacked authorized agent} in HPC. Unlike rogue agents, which authentication blocks, or adversarial agents, which authorization constrains, a hijacked agent operates with the full credentials and privileges of a legitimate user. It does not need to escalate privileges because it already has them. It does not need to bypass access controls because they already permit everything the user can do. From the perspective of every existing monitoring system that validates identity, the agent is legitimate. Yet it exfiltrates data through precisely the channel no HPC system inspects: the encrypted HTTPS connection to its LLM backend.

The attack surface is uniquely HPC. Shared parallel filesystems (e.g., Lustre) create injection vectors with no web analog, an attacker places adversarial content in a project directory, and a target agent reads it as trusted scientific data. Multi-tenant compute nodes create co-location injection: an attacker's job writes to shared \textit{/tmp}, and a co-located target job reads the poisoned content. Agent skill ecosystems create supply chain vectors: a compromised tool injects instructions from within the agent's trusted execution context. Coordinated multi-agent workflows enable distributed exfiltration that evades per-agent detection entirely.

Existing defenses are fundamentally ill-equipped against this threat. Authentication succeeds because the agent has valid Kerberos tickets. Authorization succeeds because the agent has legitimate POSIX permissions. Network monitoring is blinded because LLM API traffic is encrypted. Data loss prevention fails because sensitive data is encoded within whitelisted HTTPS requests. User behavior analytics report nothing because the agent's workflow patterns are entirely consistent with its authorized role. The defender's toolkit checks identity, but not intent.

We propose \textbf{behavioral attestation} as the foundation for securing AI agents in HPC. Our key insight is straightforward: we cannot prevent all injection attacks, because the attack surface is too large and diverse. However, we can contain their effects by detecting and responding to the behavioral violations that hijacked agents inevitably produce. When an agent accesses files outside its authorized project, connects to non-whitelisted endpoints, invokes unauthorized tools, or exceeds its data budget, attestation detects the violation and enforces containment before exfiltration occurs.

Behavioral attestation differs from prior system integrity attestation approaches along three fundamental dimensions. First, attestation provides provable guarantees rather than probabilistic alerts. Existing agent security relies on ML-based classifiers that provide alerts with inherent false positive and negative rates. Behavioral attestation provides provable guarantees: an agent either satisfies its constraints or it does not, there is no ambiguity. Second, attestation is constraint-based rather than signature-based. Existing defenses block known-bad patterns (signatures), which attackers can evade through obfuscation. Constraints define what is allowed, not what is malicious. An unauthorized action is unambiguously a violation, regardless of how it is disguised. Third, attestation operates at runtime rather than post-hoc. Existing monitoring systems alert after the attack succeeds, the damage is already done. Behavioral attestation verifies constraints continuously during execution and enforces containment before exfiltration completes.

This paper makes the following contributions. 

\begin{itemize}
    \item We formalize the hijacked authorized agent threat in HPC, identifying four distinct attack vectors that exploit unique characteristics of HPC infrastructure.
    \item We introduce behavioral attestation as a provable, constraint-based, runtime security primitive that extends zero-trust principles to autonomous AI agents.
    \item We design and implement AEGIS, a complete zero-trust architecture encompassing constraint specification, \textit{ebpf}-based attestation, policy verification, cross-node correlation for coordinated attack detection, and automated containment through Slurm integration. 
    \item We demonstrate that \projectname~achieves 100\% detection across all four attack vectors while traditional defenses achieve 0-50\%, with less than 3\% overhead on AMD EPYC hardware and zero false positives on representative workflows.
\end{itemize}

% The paper proceeds as follows. Section 2 provides background on zero-trust architecture, AI agents in HPC, HPC security, and prompt injection. Section 3 formalizes the threat model. Section 4 presents AEGIS. Section 5 describes experimental evaluation. Section 6 surveys related work. Section 7 discusses limitations and future directions. Section 8 concludes.