\section{Related Work}

\paragraph{Zero trust, workload identity, and HPC control planes.}
Zero trust shifts security policy away from implicit network trust and toward explicit authentication, authorization, and continuous policy evaluation~\cite{kindervag2010zero,rose2020zta,chandramouli2023zta}. In parallel, HPC control planes are built around scheduler-mediated resource ownership, federated identity, and shared storage semantics~\cite{neuman1994kerberos,basney2014cilogon,yoo2003slurm,lustre2025intro,ibm2024spectrumscale}. \projectname~inherits the zero-trust intuition that trust must be continuously re-established, but applies it to a different object: not only whether a user or workload has a valid identity, but whether an autonomous agent's \emph{runtime behavior} remains inside the boundary implied by its authorized task.

\paragraph{AI agents, tool use, and prompt injection.}
Recent agent research has focused on language-model-driven planning and tool use~\cite{schick2023toolformer,yao2023react,wang2023voyager,mialon2023augmented,xi2023agentsurvey,masterman2024agentsurvey}. At the same time, a growing security literature has shown that LLM-integrated applications are vulnerable to prompt injection, prompt leakage, and related control-transfer attacks~\cite{perez2022ignore,greshake2023indirect,liu2023houyi,liu2024prompt,zou2023universal,weidinger2022taxonomy,owasp2025llm}. More recent work has explored prompt-level mitigations such as Spotlighting and StruQ~\cite{hines2024spotlighting,chen2024struq}, while ToolEmu and AgentDojo provide structured environments for evaluating safety failures in tool-using agents~\cite{ruan2023toolemu,debenedetti2024agentdojo}. These works establish the threat and improve evaluation methodology, but they are primarily application- and benchmark-oriented. \projectname~instead focuses on runtime enforcement and verification for deployed agents in a shared cluster.

\paragraph{Remote attestation and runtime integrity.}
Remote attestation has long studied how to establish trust in remote software state. Sailer et al. connected attestation to access-control policy enforcement~\cite{sailer2004attestation}; Haldar et al. proposed semantic remote attestation for higher-level properties~\cite{haldar2004semantic}; PRIMA extended integrity measurement toward policy-relevant information-flow reasoning~\cite{jaeger2006prima}; and Kil et al. argued that practical attestation must reason about dynamic system properties, not just boot-time measurements~\cite{kil2009redas}. Software-based attestation such as SWATT~\cite{seshadri2004swatt} and operational systems such as Keylime and TRIGLAV illustrate the continuing move toward runtime evidence and deployable trust-management systems~\cite{schear2016keylime,ozga2021triglav}. ARI further demonstrates attestation for mission-execution integrity in real-time systems~\cite{wang2023ari}. \projectname~is complementary to this line of work, but changes the attested object: instead of only code identity or platform integrity, it attests whether an AI agent's observable actions satisfy behavioral constraints derived from its job.

\paragraph{Anomaly detection, sandboxing, and runtime monitoring.}
A long line of intrusion-detection work studies anomaly detection over audit streams, system calls, and behavior models~\cite{denning1987ids,forrest1996self,warrender1999syscalls,axelsson2000idssurvey}. Mandatory access-control and sandboxing systems such as SELinux and TrustedBSD provide stronger confinement primitives than discretionary access control alone~\cite{loscocco2001selinux,watson2003trustedbsd}. Modern runtime-security platforms such as Falco, Tetragon, and Tracee show how kernel telemetry and \ebpf can be used operationally to detect suspicious activity~\cite{falco2025docs,tetragon2025docs,tracee2025docs}. These systems are highly relevant operationally, but they are not designed to verify whether a tool-using AI agent is staying within a task-specific behavioral contract, nor do they directly reason about cross-agent covert channels in shared HPC storage.

\paragraph{Data leakage prevention.}
Data leakage prevention systems have been widely studied for data at rest, in use, and in transit. Shabtai et al. surveyed leakage-detection and prevention techniques across enterprise channels~\cite{shabtai2012dlp}, and Alneyadi et al. categorized modern DLP systems by their content- and context-analysis strategies~\cite{alneyadi2016dlp}. These systems are effective when leakage can be recognized from content signatures, protocol context, or boundary policy violations. However, they are poorly matched to the hijacked-agent setting considered here: the exfiltration channel may be a legitimate encrypted API, and the critical violation lies in the mismatch between authorized task behavior and observed action sequences. \projectname~therefore complements DLP with behavioral attestation and runtime containment rather than content inspection alone.
